\section{Challenges And Problems With Workflow Models and Management Systems}\label{commonFunc}

The review of different workflow models and management systems in
\S\ref{workflows} illustrates the diversity of solutions, the lack of a coherent
understanding of workflows per se, and the absence of a coordinated search for
higher level concepts in spite of very good past efforts. That is, there is no
``Standard Model'' that describes what a workflow is, the common elements of
workflow managements systems, or the description of how the pieces of such a
system interact to execute a workflow. Furthermore, there are few examples of
interoperability between existing systems in spite of significant community
pressure and calls for cross-system workflow execution. Poor or non-existent
interoperability is almost certainly a consequence of the ``Wild West'' state of
the field.

The state of the field does not mean that there is little or no common
functionality between workflow management systems in different domains. Many
sources in the literature, including several previously cited above, indicate
that the contrary is in fact true: there is significant duplication and
commonality in this space. The overlap in these technologies is rarely discussed
on its own merits, but instead it is commonly used to create large tables
comparing different systems, as in
\cite{ferreira_da_silva_characterization_nodate}. This creates a scenario where
more effort is expending discussing \textit{how} something is accomplished
versus the arguably more important question of \textit{what} must be
accomplished. A healthy balance of \textit{what} versus \textit{how} is
important, but we propose that the discussion of how particular problems are
solved in workflow science has overtaken the discussion of what must be
accomplished, thus creating two severe problems: \begin{itemize} \item A
``proliferation'' of tools that largely solve the same problem in the same way,
but with separate, competing implementations primarily delineated along domain,
as opposed to technological, boundaries.  \item A general lack of
interoperability and therefore inability to address larger scientific problems
using hybrid (combined) workflows, multi-facility workflow campaigns, or
heterogenous hardware without significant reimplementation.  \end{itemize}

These two problems are closely related: Tooling proliferation might not be a
problem, given sufficient resources, in the absence of calls for
interoperability between systems, and interoperability might not be an issue if
there were not so many existing systems! However, some of the most important
aspects of these problems remain separable and can be examined as such.

{\bf some suggestions to  build upon "what"} 

The primary application (workflow) needs are: lowering the burden to (i) develop, (ii)extend, and (iii) transport to another platform....platform The primary workflow system needs are: (i) determine when to use or re-use a workflow system, (ii) lower the need to develop components afresh....



\subsection{Proliferation and Common Functionality} \label{commonFunc}

The problem with the increase in the number of existing workflow management
systems have been illustrated well by reports and discussions surrounding the
future of workflow management in the Leadership Computing Facilities. The
``proliferation'' of workflow management systems and the lack of a consistent
definition of a workflow are signficant barriers to the adoption of this
technology in these facilities. The High Performance Computing Facility
Operational Assessment 2015: Oak Ridge Leadership Computing Facility (OLCF)
report, \cite{barker_scientific_2007}, describes the problem that such
facilities face, \begin{displayquote} These discussions concluded with the
observation that the current proliferation of workflow systems in response to
perceived domain-specific needs of scientific workflows makes it difficult to
choose a site-wide operational workflow manager, particularly for the
leadership-class machines. However, there are opportunities where facilities can
centralize workflow technology offerings to reduce anticipated fragmentation.
This is especially true if a facility attempts to develop, deploy, and operate
each and every workflow solution requested by the user community. Through these
evaluations, the OLCF seeks to identify interesting intersections that are of
the most value to OLCF stakeholders.  \end{displayquote} The strategy of the
OLCF is notable because it makes a very practical observation that the problem
of proliferation can be solved by consolidation of common functionality. This is
typical of an operational perspective where deployment of capability is more
important than in-depth investigation and research into how that capability
functions. 

Table \ref{blocks} lists six common types of functionality that are readily
observed in workflow management systems. There are certainly additional types of
functionality that are common, but for pedagogical reasons we limit the list to
the most obvious choices that can be discovered in a quick review of the
literature previously cited.

\begin{table*}[h] \begin{tabularx}{\textwidth}{|X|X|} \hline
\textbf{Functionality} & \textbf{Description} \tabularnewline\hline Data and
Metadata Management & Management of data, metadata, and general file input and
output activities whether for internal (tracking) or external (user)
consumption.  \tabularnewline\hline Workflow Execution Engine & The primary
actor that manages the execution of the activities as provided by the workflow
description. \tabularnewline\hline Resource Management and Acquisition &
Acquisition and management of resources, whether computing or instrumentation,
required for the successful execution of the workflow. \tabularnewline\hline
Task Management & Primary subsystem for managing individual activities, tasks or
``subworkflow'' using resources provided by the task management system. This
system is sometimes, but neither often nor exclusively, part of the Workflow
Execution Engine. \tabularnewline\hline Provenance Engine & System for tracking
execution history, sources and destinations of ingested and generated artifacts,
execution metadata including status, general logging, and provenance-based
inference tools. \tabularnewline\hline Application Programming Interface (API) &
A non-functional element of most workflow management systems that is critical to
successful deployment and maintenance of the full system as well as utilization
as a tool for creating and executing workflows. \tabularnewline\hline
\end{tabularx} \caption{Functionality commonly identified in workflow management
systems.} \label{blocks} \end{table*}

\subsection{Interoperability}

The primary driver for seeking interoperability between workflows systems is the
need to address larger scientific problems that can only be solved with
workflows that require multiple systems for complete execution. There have been
a number of community calls for interoperability. For example, Session IV of the
Twentieth Anniversary Meeting of the SOS Workshop (SOS20) focused on workflow
and workflow management system development activities of the three participating
institutions: Sandia National Laboratory, Oak Ridge National Laboratory, and the
Swiss National Supercomputing Centre, \cite{pack_sos20_2016}. Multiple
presenters illustrated the challenges facing the workflow science community and
widely agreed that no single workflow management system could satisfy all the
needs of those present. Instead, attendees proposed that the community as a
whole would be served best by seeking to enable interoperability where possible. 

Workflow interoperability however is not a simple or single attribute. There
are at least four distinct types of interoperability that merit discussion:
\begin{itemize} \item Workflow Interoperability - sharing workflows across
different science problems.  \item Execution Delegation - delegating the
execution of a workflow to a more capable or appropriate workflow management
system \item Workflow System Interoperability - executing workflows by
different workflow management systems.  \item Interchangeable Workflow System
Components - components that can be exchanged or used concurrently across one
or more systems.  \end{itemize} Two successful examples of limited
interoperability between workflow systems are discussed in
\cite{brooks_triquetrum:_2015} and \cite{mandal_integrating_2007}. Notably
both projects leveraged flavors of the Ptolemy framework, namely Triquetrum
and Kepler, and delegated the execution of workflows.

Workflow interoperability is not just a conceptual attribute, but one with
important practical implications. For example, DOE Leadership Computing
Facilities, as in \S\ref{commonFunc}, are affected by the lack of
interoperabilty (of all types). Consider the possibility that every facility
may end up supporting different workflows systems entirely such that workflows
at one facility can not be run at another without significant work to install
one or more additional workflow management systems! This idea is also
illustrated well in The Future of Scientific Workflows report through the
concept of the ``large-scale science campaign,'' \cite{deelman_future_2015}.
Such a campaign integrates multiple workflows, not necessarily all in the same
workflow management system or at the same facility, to perform data
acquisition from experimental equipment, modeling and analysis with
supercomputers, and data analysis with either grid computing or
supercomputers.


