%\section{Introduction}

Suppose for a moment there is an interesting activity that would benefit
from automation, which is known because the activity exhibits the following
properties: 
\begin{itemize} 
\item The goal of the activity is known and desirable.  
\item The tasks to achieve the goal and complete the activity are also known 
and, furthermore, are highly repetitive even in cases where decisions must be 
made to continue.  
\item The results of achieving this goal can be consumed or processed in 
standard ways. Neither magic nor clairvoyance are required.
\end{itemize}

This example may be recognized by many as a description---but not a
definition---of a workflow. Experts from many backgrounds can easily think of
activities that fit this description and even systems that automate the
activity. However, each expert will probably also imagine a different workflow:
A businessperson might imagine the workflow for processing payments; a medical
professional might imagine updating medical charts and records; and
scientists might imagine performing an analysis with modeling and simulation
software, analyzing a large amount of data, or quantifying uncertainty.
Within the scientific community this has led to a rather predictable situation:
Everyone has a different definition of workflow and has created his or her own
systems for managing and processing workflows.

This leads to some practical consequences for scientific
workflows. In spite of the similarities in high-level abstractions and
higher-order concepts, extremely specialized software solutions have been
developed and communities have formed to process scientific workflows. These
differences hold across scientific problems, with each generally providing some
level of service that was not or perhaps is not available in a regular programming
language, system library, or problem-solving workbench. These systems have
accreted workflow management capabilities over time that have effectively
resulted in the creation of large, monolithic software stacks that cannot
communicate among each other, require a large amount of expertise to
use, often put high demands on back-end systems either by design or
through assumptions, and are often too specialized to jump between workflow
execution for data analysis and modeling and simulation. 

Recent developments suggest that this situation is not ideal and could be
signficantly costly to maintain in the future. The continued scalability,
sophistication, and maintainability of large, monolithic systems is called into
question as scientific problems become more complex, functionality moves from
libraries to operating systems, and open source development continues to rise
as the dominant means of collaborating on software. Software complexity, in
particular, often makes it impossible for development on large systems to scale
to the required level because the accretion of new capabilities means managing
larger pools of people and a larger development effort. One obvious
alternative with some degree of historical precedence in the field is to
develop common building blocks that provide common services used to both define
and execute workflows. The astute reader may ask ``Why has this alternative
failed in the past?'' The answer to that question lies in the very nature of
workflow science's diverse ecosystem. 
%Such an approach not only makes it possible to coalesce
%around a standard definition and understanding of workflows, but also make it
%possible to separate and distribute the work required to construct the building
%blocks from the effort to define workflows and to create workflow management
%systems that may share the building blocks while retaining required
%customizations. This article contributes to the ongoing discussion by providing
%\begin{itemize}
%\item an illustration of the diverse nature of
%scientific workflows (\S \ref{workflows}) that describes the different
%areas where scientific workflows and systems have appeared in the literature,
%how they have been classified in the past, and the arguments around coalescence
%that are driven by calls of interoperability (\S\ref{interop} and
% \S\ref{commonFunc});
%\item a description of the necessary
%subset of functionality that is common across a number of scientific workflow
%management systems that would, in principle, be good candidates for
%consolidation and sharing (\S \ref{buildings-blocks}); and
%\item an understanding of
%these common elements as building blocks and how composing these building
% blocks addresses a number of the problems not easily addressed by the
% monolithic design
%of existing systems (\S\ref{buildings-blocks} and \S\ref{discussion}).
%\end{itemize}

