%%This is a very basic article template.  %%There is just one section and two
%%subsections.  
\documentclass[10pt,conference,final]{IEEEtran}

\usepackage{cite} 
\usepackage{tabularx}
\usepackage[autostyle]{csquotes} 
\usepackage{url} 
\usepackage{todonotes}
\usepackage{float}
\usepackage{array}
\usepackage{tabu}

\begin{document}

\title{Toward Common Components for Open Workflow Systems}


\author{\IEEEauthorblockN{Jay Jay Billings} \IEEEauthorblockA{Oak Ridge National
Laboratory\\ PO Box 2008 MS 6173\\ Oak Ridge, TN USA 37831\\ Email:
billingsjj@ornl.gov\\ Twitter: @jayjaybillings} \and \IEEEauthorblockN{Shantenu
Jha} \IEEEauthorblockA{RADICAL, ECE\\ Rutgers University, \\ Piscataway, NJ USA
08854\\ Email: shantenu.jha@rutgers.edu} }

\maketitle

\begin{abstract}

The role of scalable high-performance workflows and flexible workflow
management systems that can support multiple simulations will continue to
increase in importance. For example, with the end of Dennard scaling, there is
a need to substitute a single long running simulation with multiple repeats of
shorter simulations, or concurrent replicas. Further, many scientific problems
involve many simulations to support order to solve a higher-level problem or
statistically meaningful science. However most supercomputing software
development and performance enhancements have focused on optimizing single-
simulation performance. On the other hand, there is a strong inconsistency in
the definition and practice of workflows and workflow management systems. This
inconsistency often centers around the difference between several different
types of workflows, including modeling and simulation, uncertainty
quantification, and purely conceptual workflows. This work explores this
phenomenon by examining the different types of workflows and workflow
management systems, reviewing the perspective of a large supercomputing
facility, examining the common features and problems of workflow management
systems, and finally presenting a proposed solution based on the concept of
common building blocks. The implications of the continuing proliferation of
workflow management systems and the lack of interoperability between these
systems are discussed from a practical perspective. In doing so, we have begun
an investigation of the design and implementation of open workflow systems for
supercomputers based upon common components.

% based approach and thereby open workflow systems for of workflow systems for
% supercomputing.

% The study of “workflow science” has an open secret: Discussions about workflows and workflow management systems inevitably turn into discussions about the nature of those very topics! Too often, surveying a number of papers will result in different definitions from each paper and a strong inconsistency that leaves the reader with a skewed opinion. This inconsistency often centers around the different between several different types of workflows, including grid, modeling and simulation, uncertainty quantification, and purely conceptual workflows. This work explores this phenomenon by examining the different types of workflows and workflow management systems, reviewing the perspective of a large supercomputing facility, examining the common features and problems of workflow management systems, and finally presenting a proposed solution based on the concept of common building blocks. The implications of the continuing proliferation of workflow management systems and the lack of interoperability between these systems are discussed from a practical perspective.

\end{abstract}
\hfill \\
\underline{Notice of Copyright:} This manuscript has been authored by UT-
Battelle, LLC under Contract No. DEAC05-00OR22725 with the U.S. Department of
Energy. The United States Government retains and the publisher, by accepting
the article for publication, acknowledges that the United States Government
retains a nonexclusive, paid-up, irrevocable, world-wide license to publish or
reproduce the published form of this manuscript, or allow others to do so, for
United States Government purposes. The Department of Energy will provide
public access to these results of federally sponsored research in accordance
with the DOE Public Access Plan (http://energy.gov/downloads/doe-public-access-plan).



\input{introduction}
\input{workflows-review}
\input{experience}
\input{common}
\input{buildingblocks}
\input{discussion}
\input{ack}

% The bibliography 
\bibliographystyle{IEEEtran}
\bibliography{IEEEabrv,bib}

\end{document}
